---
title: "R Notebook"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

## Database preparation 

The first step is to prepare the database and clean it

```{r}
library(haven)
library(survey)
CMS_DATA_BG <- read_sav("CMS_logical_corrections_V1.sav")
head(CMS_DATA_BG)
```

```{r}
library(dplyr)
CMS_DATA_BG <- CMS_DATA_BG %>%
  mutate(
    # keep existing Changed_dummy entries; treat NA as 0 only for this update step
    Changed_dummy = if_else(is.na(Changed_dummy), 0, Changed_dummy),

    missing_key = if_else(
      is.na(D2) | D2 >= 99 |                  # Age
      is.na(D1) | !(D1 %in% c(1, 2)) |        # Sex
      is.na(D26) | !(D26 %in% c(1, 2, 3, 4)), # Settlement
      1, 0
    ),

    # preserve prior 1s, add new 1s where key covariates are bad
    Changed_dummy = if_else(Changed_dummy == 1 | missing_key == 1, 1, 0)
  )

df_clean <- CMS_DATA_BG %>%
  mutate(
    age_group = case_when(
      D2 >= 18 & D2 <= 24 ~ "18-24",
      D2 <= 34 ~ "25-34",
      D2 <= 44 ~ "35-44",
      D2 <= 54 ~ "45-54",
      D2 <= 64 ~ "55-64",
      D2 >= 65 ~ "65+"
    ),
    sex = factor(D1, levels = c(1, 2), labels = c("Male", "Female")),
    urban = case_when(
      D26 %in% c(1, 2) ~ "City",
      D26 == 3 ~ "Town",
      D26 == 4 ~ "Rural"
    )
  )
```

```{r}
# Share dropped
table(df_clean$Changed_dummy)

# Distributions
table(df_clean$age_group)
table(df_clean$sex)
table(df_clean$urban)
```

## Raking and post-strat weights construction

```{r}
library(eurostat)
library(dplyr)
library(lubridate)
pop_raw <- get_eurostat("demo_pjan", time_format = "num")

bg_pop <- pop_raw %>%
  filter(
    geo == "BG",
   TIME_PERIOD== 2025,              # survey year
    sex %in% c("M", "F"),
age %in% c(paste0("Y", 18:99), "Y_GE90")
  )
```

```{r}
bg_pop_age_sex <- bg_pop %>%
  mutate(
    age_num = case_when(
      age == "Y_GE90" ~ 90,
      TRUE ~ as.numeric(sub("^Y", "", age))
    ),
    sex = recode(sex,
      "M" = "Male",
      "F" = "Female"
    ),
    age_group = case_when(
      age_num >= 18 & age_num <= 24 ~ "18-24",
      age_num >= 25 & age_num <= 34 ~ "25-34",
      age_num >= 35 & age_num <= 44 ~ "35-44",
      age_num >= 45 & age_num <= 54 ~ "45-54",
      age_num >= 55 & age_num <= 64 ~ "55-64",
      age_num >= 65               ~ "65+"
    )
  ) %>%
  filter(!is.na(age_group)) %>%
  group_by(age_group, sex) %>%
  summarise(pop = sum(values, na.rm = TRUE), .groups = "drop") %>%
  mutate(Freq = pop / sum(pop))

bg_pop_age_sex
```

```{r}
pop_sex <- bg_pop_age_sex %>%
  group_by(sex) %>%
  summarise(pop = sum(pop), .groups = "drop") %>%
  mutate(Freq = pop / sum(pop))

pop_sex

sum(bg_pop_age_sex$Freq)

```

```{r}
urb_raw <- get_eurostat("ilc_lvho01", time_format = "num")

#https://ec.europa.eu/eurostat/databrowser/view/ILC_LVHO01__custom_1762022/default/table?lang=en

unique(urb_raw$deg_urb)
bg_urban <- urb_raw %>%
  filter(
    geo == "BG",
    TIME_PERIOD == 2025,
    deg_urb %in% c("DEG1", "DEG2", "DEG3"), 
    incgrp=="TOTAL", 
    building=="TOTAL"
  )
```

```{r}
pop_urban <- bg_urban %>%
  mutate(
    urban = case_when(
      deg_urb == "DEG1" ~ "City",
      deg_urb == "DEG2" ~ "Town",
      deg_urb == "DEG3" ~ "Rural"
    ),
    Freq = values / 100
  ) %>%
  select(urban, Freq)
sum(pop_urban$Freq)
```

```{r}
bg_pop_age_sex <- bg_pop_age_sex %>%
  mutate(
    age_group = factor(age_group,
      levels = c("18-24", "25-34", "35-44", "45-54", "55-64", "65+")),
    sex = factor(sex,
      levels = c("Male", "Female"))
  )

pop_urban <- pop_urban %>%
  mutate(
    urban = factor(urban,
      levels = c("City", "Town", "Rural"))
  )
```

```{r}
design <- svydesign(
  ids = ~1,
  data = df_clean
)
raked <- rake(
  design,
  sample.margins = list(~age_group + sex, ~urban),
  population.margins = list(
    bg_pop_age_sex %>% select(age_group, sex, Freq),
    pop_urban %>% select(urban, Freq)
  )
)
df_clean$weight <- weights(raked, type = "analysis")
summary(df_clean$weight)

df_clean$weight_ipw <- 1 / raked$prob
summary(df_clean$weight_ipw)
```

```{r}
df_clean$weight_final <- df_clean$weight / mean(df_clean$weight)
summary(df_clean$weight_final)
write_sav(df_clean, "weighted_df.sav")

```

```{r}
# Urban
svytable(~urban, raked) / sum(svytable(~urban, raked))

# Age × Sex
svytable(~age_group + sex, raked) / sum(svytable(~age_group + sex, raked))
```

###Финална база данни в теглен дизайн

```{r}
design_w <- svydesign(
  ids = ~1,
  data = df_clean,
  weights = ~weight_final
)
```

Сега ще тестваме някакви неща:

```{r}
svymean(~A2, design_w, na.rm = TRUE)
```

```{r}
svytable(~sex, design_w)
prop.table(svytable(~urban, design_w))
```

```{r}
prop.table(svytable(~urban + sex, design_w), margin = 1)
```

```{r}
summary(svyglm(A2 ~ sex + age_group + urban, design = design_w))
```

# Stoyanov's indexes

| SCI Indicator | Theoretical Mechanism | Survey Question(s) |
|:----------------------:|:----------------------:|:----------------------:|
| **L1: Consultation Exclusion** | Bypassing public scrutiny through fast-tracking. | **L_PROC (Item 1):** Frequency of accelerated procedures to avoid debate. **SC5:** Perception of whose interests prevail in decisions. |
| **L2: RIA Manipulation** | Use of formalistic or fake evidence to justify laws. | **L_PROC (Item 2):** Formalism of impact assessments. **L_PROC (Item 3):** Direct authorship by lobbyists. |
| **L3: Transparency Resistance** | Refusal to regulate lobbying and executive conflicts. | **L_LOBBY:** Deliberate avoidance of lobbying regulations. **A9/A10:** Social acceptability of favors for MPs/Ministers. |

## Consultancy exclusion

Since **Consultation Exclusion** is constructed from `L_PROC_A` and `SC5`, and one variable is ordinal while the other is categorical, the two indicators must be normalized to make them comparable in scale.

First, `SC5` is recoded into a binary indicator of exclusion from the public interest:

$$
sc5\_excl=\mathbb{1}\{SC5\neq 1\}
$$

where `SC5 = 1` corresponds to “The public”, and all other categories indicate that some non-public interest is perceived to prevail.

Second, because `L_PROC_A` is ordinal, it is transformed to the unit interval:

$$
l\_proc\_{01}=\frac{4-L\_PROC\_A}{3}
$$

This transformation reverses the original coding so that higher values indicate more exclusion. Thus, if `L_PROC_A = 1`, this maps to 1, while if `L_PROC_A = 4`, this maps to 0.

Finally, the **Consultation Exclusion** index is calculated as the simple average of the two normalized components:

$$
consultation\_exclusion=\frac{l\_proc\_{01}+sc5\_excl}{2}
$$

```{r}
design_w <- update(
  design_w,
  sc5_excl = ifelse(
    SC5 == 1, 0,
    ifelse(SC5 %in% c(2, 3, 4, 5), 1, NA)
  ),
  l_proc_01 = ifelse(L_PROC_A %in% 1:4, (4 - L_PROC_A) / 3, NA)
)

design_w$variables$consultation_exclusion <- ifelse(
  complete.cases(design_w$variables[, c("l_proc_01", "sc5_excl")]),
  rowMeans(design_w$variables[, c("l_proc_01", "sc5_excl")]),
  NA
)

svymean(~consultation_exclusion, design_w, na.rm = TRUE)
svyquantile(
  ~consultation_exclusion,
  design_w,
  quantiles = c(0, 0.25, 0.5, 0.75, 1),
  na.rm = TRUE
)
```

```{r}
svytable(~SC5, design_w)
svytable(~L_PROC_A, design_w)
svymean(~sc5_excl + l_proc_01, design_w, na.rm = TRUE)
```

The results indicate a **systematically high level of consultation exclusion** in the legislative process. Survey evidence shows that respondents overwhelmingly perceive that:

-   Accelerated procedures are frequently used, limiting debate and scrutiny; and

-   Public interest is rarely the primary driver of decisions, with non-public actors, particularly business and political interests, are seen as dominant.

## **RIA Manipulation** 

```{r}
design_w <- update(
  design_w,
  l_proc2_01 = ifelse(L_PROC_B %in% 1:4, (4 - L_PROC_B) / 3, NA),
  l_proc3_01 = ifelse(L_PROC_C %in% 1:4, (4 - L_PROC_C) / 3, NA)
)

design_w$variables$ria_manipulation <- ifelse(
  complete.cases(design_w$variables[, c("l_proc2_01", "l_proc3_01")]),
  rowMeans(design_w$variables[, c("l_proc2_01", "l_proc3_01")]),
  NA
)
svymean(~ria_manipulation, design_w, na.rm = TRUE)

svyquantile(
  ~ria_manipulation,
  design_w,
  quantiles = c(0, 0.25, 0.5, 0.75, 1),
  na.rm = TRUE
)
```

The results indicate a consistently high perception that Regulatory Impact Assessments are used in a formalist manner and do not function as genuine evidence-based tools. At the same time, respondents report that legislative drafting is often influenced or directly carried out by external actors. Taken together, these findings suggest that RIA processes are not serving as safeguards against capture, but rather as instruments that legitimize decisions shaped outside formal institutional channels.

## **Transparency Resistance**

```{r}
design_w <- update(
  design_w,
  l_lobby_01 = ifelse(L_LOB %in% 1:4, (4 - L_LOB) / 3, NA)
)
```

```{r}
design_w <- update(
  design_w,
  a9a_01  = ifelse(A9A  %in% 1:4, (4 - A9A)  / 3, NA),
  a9b_01  = ifelse(A9B  %in% 1:4, (4 - A9B)  / 3, NA),
  a9c_01  = ifelse(A9C  %in% 1:4, (4 - A9C)  / 3, NA),
  a9d_01  = ifelse(A9D  %in% 1:4, (4 - A9D)  / 3, NA),

  a10a_01 = ifelse(A10A %in% 1:4, (4 - A10A) / 3, NA),
  a10b_01 = ifelse(A10B %in% 1:4, (4 - A10B) / 3, NA),
  a10c_01 = ifelse(A10C %in% 1:4, (4 - A10C) / 3, NA),
  a10d_01 = ifelse(A10D %in% 1:4, (4 - A10D) / 3, NA)
)

svymean(~A9A + a9a_01, design_w, na.rm = TRUE)

```

```{r}
design_w$variables$transparency<- rowMeans(
  design_w$variables[, c(
    "a9a_01","a9b_01","a9c_01","a9d_01",
    "a10a_01","a10b_01","a10c_01","a10d_01"
  )],
  na.rm = TRUE
)
```

```{r}
svymean(~transparency, design_w, na.rm = TRUE)
```

```{r}
vars <- design_w$variables[, c(
  "consultation_exclusion",
  "ria_manipulation",
  "transparency", 
  "l_lobby_01"
)]

cor(vars, use = "complete.obs")
```

```{r}
vars_clean <- na.omit(vars)
fa <- factanal(
  vars_clean,
  factors = 1,
  scores = "regression"
)

fa
```

```{r}
vars_clean<-vars_clean[-3]
fa <- factanal(
  vars_clean,
  factors = 1,
  scores = "regression"
)

fa
```
